Adult zebrafish telencephalon and optic tectum serve different behavioral and circuit functions. If proteoforms matter to that division of labor, the difference should appear in a region-resolved molecular map. We asked whether the public top-down proteomics record already shows such a difference. It does. Using article-level counts together with the released Tel2/Teo2 proteoform tables from a 2022 laser-capture microdissection-capillary zone electrophoresis-tandem mass spectrometry study, we find that the two regions occupy sharply distinct proteoform regimes rather than lightly shifted versions of the same local proteome. The article reports 35 shared proteoforms out of a union of 807 (Jaccard $=0.0434$); the released exact accession+proteoform tables expose 29 shared entries out of 805 (exact-ID Jaccard $=0.0360$, S{\o}rensen $=0.0689$, weighted Jaccard $=0.0427$). Canonicalized accession-plus-sequence matching raises the shared count to 44, placing the article's 35 between strict and relaxed table-matching rules, while abundance-based similarities remain low under per-run total-intensity, upper-quartile, and median-ratio normalization (weighted Jaccard $=0.0430$--$0.0557$, Bray--Curtis $=0.0820$--$0.1055$). A prevalence-adjusted lower-tail overlap test points in the same direction: the exact-ID centered Jaccard is $-0.2792$ relative to the independence baseline, with $p = 5.56\times10^{-183}$, and even the protein-collapsed overlap remains far below marginal-prevalence expectation ($p = 5.84\times10^{-33}$). Duplicate-informed Chao2 and jackknife sensitivity analyses still leave the overlap in the same low range (Jaccard $=0.0233$--$0.0394$ for Chao2 scenarios; $0.0272$ for jackknife), and a minimal MNAR-style low-intensity fill only shifts weighted Jaccard to $0.0456$--$0.0483$. The separation is not merely sparse; it is biologically ordered. Telencephalon-associated markers contribute 23 matched counts versus 2 spillover counts, whereas optic-tectum-associated markers contribute 102 matched counts versus 1 spillover count, corresponding to a marker alignment fraction of 0.9766 and a directional exact-test odds ratio of 1173 (one-sided $p = 5.12\times10^{-22}$). A family-size-preserving randomization never reproduced that alignment in 200{,}000 trials ($p < 5\times10^{-6}$). A separate literature-derived composition panel built from adult zebrafish myelin markers and adult optic-tectum radial-glia markers also aligns strongly (32 matched counts versus 3 spillover counts; odds ratio $=73.3$, one-sided $p = 6.68\times10^{-4}$). The N-terminal acetylation contrast proves less stable: a coarse logistic sensitivity adjusting for precursor mass, sequence span, and first-residue position yields an odds ratio of 2.94 (95\% CI 0.70--12.30; $p = 0.140$), and the first-residue$\leq2$ subset yields no enrichment. Acetylation is therefore better treated as hypothesis-generating than as a regional signature. The conclusion is plain. Telencephalon and optic tectum differ strongly at the proteoform level, and the difference follows the known biology of the two regions.
