\section{Introduction}

Region-resolved proteomics is most informative when it can expose molecular differences that parallel functional anatomy. In adult zebrafish, the telencephalon is implicated in forebrain functions including social orienting, learning, memory, and other higher-order behavioral regulation \citep{stednitz2018social,reemst2023memory,pandey2023telencephalon}. The optic tectum, by contrast, is a classic visuomotor structure that transforms sensory information into orienting and avoidance actions \citep{roeser2003tectum,orger2016visuomotor}. If those regions maintain distinct proteoform repertoires, a region-resolved top-down proteomics record should be able to detect that difference.

The 2022 LCM-CZE-MS/MS study of \citet{lubeckyj2022lcmcze} offers a useful test. It isolated adult zebrafish telencephalon and optic tectum and identified hundreds of proteoforms from each region using a top-down workflow whose downstream interpretation would normally depend on tools such as TopPIC and related proteoform-level post-processing \citep{choi2022toppic}. The article argued that the resulting profiles matched basic regional function, but the evidence for that interpretation remained dispersed across the text and figures. The accompanying PRIDE release exposes region-specific Tel2 and Teo2 proteoform tables, making it possible to examine exact accession+proteoform overlaps, duplicate incidence patterns, marker-family membership, and simple intensity summaries without pretending to have rerun the full raw-feature pipeline.

The question addressed here is therefore straightforward: what does this released top-down dataset already establish about regional molecular organization in the adult zebrafish brain? Read quantitatively, it supports a sharply separated telencephalon-versus-tectum proteoform architecture that also tracks known regional biology. Just as important, the released duplicate structure lets us ask whether that separation could be explained away by uneven recovery or identifier noise. Within the limits of the public tables, it cannot. Prior zebrafish functional studies and atlases tell us what the two regions are generally for \citep{stednitz2018social,pandey2023telencephalon,roeser2003tectum,orger2016visuomotor}, while recent top-down zebrafish-brain studies show that proteoform-level structure remains biologically informative well beyond this original pilot \citep{askarani2025aging}. In parallel, spatial-omics foundation models such as KRONOS and HEIST emphasize large-scale representation learning across imaging or count-based spatial assays \citep{shaban2025kronos,madhu2025heist}. What has been missing is a compact proteoform-specific account that states, in quantitative terms, what this public dataset already implies about regional molecular organization.

Three results follow. First, the released telencephalon and optic-tectum tables imply a much sharper proteoform-level separation than a casual reading of the source paper would suggest, and that result is visible across article-level, exact-ID, canonicalized, protein-collapsed, prevalence-adjusted, and duplicate-based detectability views. Second, the separation is biologically ordered rather than merely sparse: once the source paper's marker prose is rewritten as a matched-versus-spillover problem, the signal remains strong under exact testing, interval estimation, family-size-preserving randomization, protein collapse, and an independent literature-derived composition panel that is not inherited from the source paper. Third, the visible acetylation asymmetry does not survive simple detectability-aware follow-up and is better treated as hypothesis-generating than as a stable regional program. Throughout, the analysis is packaged with code, figures, exact marker-family membership tables, standards-aware canonicalization examples, and machine-readable evidence traceability so that the argument can be checked locally rather than taken on trust.
