\section{Data and Analysis Setup}

\subsection{Source material}

The quantitative core of this paper comes from \citet{lubeckyj2022lcmcze}, a pilot study of adult zebrafish telencephalon and optic tectum using laser capture microdissection coupled to capillary zone electrophoresis tandem mass spectrometry. We use three public source layers. The first is the article itself, from which we extract region totals, protein overlap, named marker proteoforms, technical replicate summaries, and post-translational modification counts into a local evidence file. The second is the journal supplementary workbook, whose note sheet states that the released proteoforms were identified in TopPIC with a 1\% proteoform-spectrum-match FDR and a 5\% proteoform-level FDR, and that duplicate label-free quantification was performed with TopDiff. The third is the released PRIDE region tables \texttt{Proteoform\_ID\_Tel2.xlsx} and \texttt{proteoform\_ID\_Teo2.xlsx}, which expose machine-readable accession-level proteoform lists, duplicate match flags, and per-run intensity columns. The supplement's \texttt{Tel2} and \texttt{Teo2} sheets match those standalone PRIDE tables exactly, so we treat the supplementary workbook as provenance metadata rather than as an additional quantitative dataset. We parse the public tables directly with a local standard-library reader and preserve exact row-level provenance in a source-table CSV bundle, a marker-membership CSV, and a machine-readable traceability JSON. The OmicsDI landing page also lists the four raw duplicate files (\texttt{Tel2\_1.raw}, \texttt{Tel2\_2.raw}, \texttt{Teo2\_1.raw}, and \texttt{Teo2\_2.raw}); we record those links in the manifest, but we do not attempt raw-file reprocessing in this note because the processed TopPIC feature histories needed for a faithful reanalysis are not part of the public release.

For exact-ID overlap we treat the biological unit as the accession-plus-proteoform pair. This is stricter than comparing only proteoform strings, because the same peptide-like string can occur under more than one accession in the released tables. For the curated marker panel, we use transparent matching rules tied to the source paper's own language: exact gene-symbol matches for NPY, PENKB, PYY-A, and neurogranin a, and description-keyword matches for synucleins, reticulon, myosin, actin, and troponin. Those rules are written into the local traceability files so that a reader can inspect or contest them directly.

\subsection{Derived metrics}

The local computation produces six classes of outputs.

First, regional separation is summarized at three linked levels. The article-reported regional split is described by the published Jaccard overlap
\[
\Jaccard = \frac{|\Tel \cap \Teo|}{|\Tel \cup \Teo|}.
\]
The released source tables then support an exact-ID Jaccard, an exact-ID S{\o}rensen overlap, a weighted Jaccard based on average duplicate intensities, and a Bray--Curtis similarity on the same intensity vectors. To make the 35-versus-29 shared-proteoform discrepancy more auditable, we also compare strict accession+proteoform matching with a relaxed accession-plus-canonicalized-sequence match that strips bracketed PTM annotations and punctuation from the proteoform string. Reporting all three views is important because the article aggregate and the released tables are biologically consistent but not numerically identical. We complement those point estimates with a 10{,}000-draw row-resampling sensitivity analysis for exact-ID, canonicalized, and protein-collapsed overlaps.

To answer the ``low overlap relative to what?'' criticism more directly, we also report a prevalence-adjusted overlap significance check inspired by \citet{chung2019jaccard}. With the observed telencephalon and optic-tectum margins held fixed, the overlap count is tested against the fixed-margin hypergeometric null, which is monotone with the Jaccard coefficient for a given pair of margins. We summarize that comparison by the observed Jaccard, the marginal-prevalence independence baseline, the centered Jaccard (observed minus expected), and the exact lower-tail $p$ value.

For the abundance-weighted comparison, each proteoform receives the mean of its two duplicate intensities within a region, blank cells are treated as zero, and no log transform is applied. If $x_i$ and $y_i$ are the resulting region-level intensities for exact ID $i$, then the weighted Jaccard is
\[
J_w(x,y)=\frac{\sum_i \min(x_i,y_i)}{\sum_i \max(x_i,y_i)},
\]
and the Bray--Curtis similarity is
\[
BC(x,y)=1-\frac{\sum_i |x_i-y_i|}{\sum_i (x_i+y_i)}.
\]
We report both point estimates and row-bootstrap sensitivity intervals for these intensity-based similarities.

Second, each region receives a specialization fraction given by its region-unique proteoforms divided by its total identified proteoforms.

Third, for each marker group we compute a stabilized bias score
\[
\log_2 \frac{\Tel + 1}{\Teo + 1},
\]
which is positive for telencephalon-biased markers and negative for optic-tectum-biased markers.

Fourth, we aggregate the curated markers into two functional axes and summarize matched-region concentration, Wilson confidence intervals for the axis-level alignment fractions, an approximate log-odds confidence interval for the $2\times2$ matched-versus-spillover table, and an exact directional test on the same table. We repeat that concentration check after collapsing proteoforms to unique proteins, and we compute an intensity-weighted matched-region fraction using the mean of the two run intensities in each released source-table row. To address circularity concerns more directly, we also run a family-size-preserving randomization test that keeps the nine marker-family sizes and the global 25-versus-103 telencephalon/optic-tectum label totals fixed while shuffling region labels across the 128 marker-panel counts. The exact test is directional because the biological hypothesis is directional: the telencephalon-associated markers should concentrate in telencephalon and the optic-tectum-associated markers should concentrate in optic tectum rather than the reverse. In addition, we curate a second, explicitly independent composition-oriented panel that is not inherited from the source paper: adult zebrafish CNS myelin markers (PLP1b, MBPa, MBPb, MPZ/P0) from the zebrafish myelin proteome literature \citep{siems2021myelin}, and adult optic-tectum radial-glia markers (GFAP, FABP7a, S100B) from the adult tectal precursor literature \citep{ito2010tectum}. Those exact-gene rules are applied to the same released source tables and scored with the same alignment statistics.

Fifth, we run reviewer-motivated missingness, normalization, contamination, and notation-sensitivity checks that are feasible from the public data. One-to-three matched counts per functional axis are pessimistically reassigned to spillover, and the motor-associated myosin, actin, and troponin families are excluded entirely to probe whether the optic-tectum signal could be carried mainly by a local contamination story. We also screen a small sentinel panel of glial-support markers (GFAP, S100B), myelin markers (MBPA, MBPB), housekeeping markers (EEF1A, GAPDH-2), and structural markers (TUBA1C, TUBB2B, NEFMA) from the same released tables as a lightweight purity guardrail. For abundance similarity, we compare the raw mean-duplicate intensities with three identifiable alternatives: per-run total-intensity scaling before averaging duplicate rows, region-level total-sum scaling, upper-quartile scaling, and median-ratio scaling on shared exact IDs. We also report the four explicit duplicate run-pair similarities so the reader can see how much the abundance story shifts from run to run before averaging. Blank intensity cells are treated as absent detections in that run and therefore contribute zero intensity; no imputation or pseudo-counts are added in the baseline analysis. Because the review specifically asked for a minimal MNAR-aware check, we add two low-intensity-limit sensitivities in which undetected exact IDs are imputed to half of the region-specific 5th-percentile or 10th-percentile nonzero intensity. With only two duplicate runs per region, richer SVD-based normalization or missing-value-imputation frameworks are better treated as future-work comparators than as estimable models in the present package \citep{karpievitch2009eigenms,karpievitch2012missing,lazar2016missing}. For missingness we use the duplicate \texttt{Match}/\texttt{No match} flags in the released Tel2 and Teo2 tables to compute the incidence counts $q_1$ and $q_2$, then form both Chao2 and first-order incidence jackknife lower-bound richness estimates \citep{chao2005similarity}. Higher-order iChao-style corrections are not identifiable from the public tables because only the two-run duplicate counts are exposed, not the higher-frequency incidence counts required for those estimators \citep{chiu2014goodturing}. Because the public tables do not expose cross-region duplicate co-detection directly, we report a sensitivity envelope: one version holds the observed shared overlap fixed while richness increases, and others scale the shared overlap by the smaller, larger, or geometric-mean duplicate-inflation factors. We also compare three deterministic relaxed matching rules beyond the strict exact-ID definition so that the discrepancy diagnostic is not resting on a single stripping heuristic.

Sixth, we summarize the post-translational modification inventory and the technical sensitivity values reported for the duplicate analyses and the single-run high-water mark. Within the matched marker panel, we test whether N-terminal acetylation is regionally biased by comparing the acetylated-versus-non-acetylated counts in the telencephalon-matched and optic-tectum-matched marker subsets using the same one-sided exact-test logic as above. To answer the multiple-testing criticism directly, we also run one exploratory companion screen in which all non-acetyl bracketed mass shifts are collapsed into a single ``other mass shift'' bucket and then apply a Holm correction across the two PTM screens. That makes acetylation the only pre-specified, source-paper-driven PTM comparison and keeps the exploratory mass-shift screen clearly separated from it. To probe detectability confounding more directly, we add two reviewer-driven follow-ups based on the public-table covariates that actually exist: permutation tests for precursor mass, sequence span, first-residue position, and mean intensity among the acetylated matched subsets, plus a coarse logistic sensitivity in which region label is modeled as a function of acetylation, log precursor mass, sequence span, and first-residue position. We also report a restricted recheck on the subset whose first residue is at or within two positions of the protein N terminus.

\subsection{Interpretive scope}

The analysis stays close to what the public release can genuinely support. It does not estimate significance across the full proteome, recover the raw TopPIC feature tables, or reinterpret unidentified mass shifts. The strongest function-alignment claims therefore remain tied to a curated marker panel taken directly from the source article's biological discussion. What the released region tables and supplementary note do provide, however, is enough structure to go well beyond article-only extraction: exact-ID overlaps, duplicate-informed missingness sensitivity, intensity-weighted marker concentration, protein-collapsed robustness, run-pair abundance stability, region-linked PTM summaries, and explicit table-side threshold metadata. Those additions make the central biological claim much easier to inspect, even though the deeper feature-level quantities from the raw TopPIC workflow --- FDR histories, localization confidence, and replicate-resolved cross-region incidence --- are still unavailable \citep{choi2022toppic}. The aim throughout is simply to say as clearly as possible what the public paper, supplement, and released tables already establish about adult zebrafish regional proteoform organization.
