Gleason's original theorem is deep, important, and notoriously hard to explain to beginners. This note does not claim a new elementary proof of the original 1957 theorem. Instead, it isolates a simpler target that still captures the density-matrix conclusion in finite dimension. We prove a finite-dimensional real Gleason-type theorem for additive rules on the effect space of symmetric matrices. The proof is built in three steps. First, we prove an interval lemma: every bounded additive rule on $[0,1]$ is linear. Second, we turn that lemma into a high-school-friendly sphere corollary for rules that depend only on squared height. Third, we lift the same additivity idea to real symmetric matrices and show that every normalized additive rule on effects has the form $E \mapsto \mathrm{tr}(\rho E)$ for a unique positive semidefinite trace-one matrix $\rho$. The note avoids quantum-mechanics vocabulary and keeps the proof in bounded additivity, orthogonality, and finite-dimensional linear algebra. It also states the gap to the original theorem explicitly: the argument is finite-dimensional, real, and assumes additivity on the full effect space rather than only on orthogonal projections. To satisfy the repository's reproducibility standard, we also ship generated numerical sanity checks and high-resolution figure assets. The computations do not replace the proof, but they do confirm that the reconstruction formulas behave at machine precision on random sampled examples. The result should therefore be read as an entry point into Gleason-style rigidity rather than as a substitute for the classical theorem.
