\section{From Workaround to Interface}

\input{figures/TABLE_source_vs_upgrade.tex}
\input{figures/FIG_serialization_view.tex}

The central reframing of this manuscript is straightforward: serialization should be treated as a modeling interface rather than as a preprocessing workaround. This section explains why that reframing is justified by the literature that appeared after the source paper.

The original paper demonstrated that a Hilbert-style traversal could map voxelized 3D structures into lower-dimensional arrays while preserving enough local organization for CNN classification \citep{corcoran2018spatial}. At the time, the practical benefit was obvious. Volumetric models were expensive, low-resolution, and poorly matched to high-channel structural data. A reversible mapping into 2D or 1D offered a way to borrow mature lower-dimensional architectures without discarding all spatial structure.

If the field had simply moved toward stronger native 3D models, the 2018 story might have aged into a historical curiosity. But that is not what happened. Certainly, native point-based models such as Point Transformer raised the baseline for direct geometric learning \citep{zhao2021pointtransformer}. Yet later work also made serialization itself more explicit. Point Transformer V3 replaced exact neighborhood operations with efficient serialized neighbor mappings to improve scale \citep{wu2024ptv3}. PointMamba used space-filling curves for point tokenization in a linear-complexity state-space model \citep{liang2024pointmamba}. Voxel Mamba treated serialization as the basis for group-free sequence modeling in 3D detection \citep{zhang2024voxelmamba}. NoKSR leveraged serialized orderings as a practical approximation layer for neighborhood retrieval in surface reconstruction \citep{li2025noksr}. Across these settings, ordering is no longer incidental. It actively shapes what computation is feasible.

This is why the source paper deserves reinterpretation. Its durable contribution is not that 2D CNNs once matched a particular 3D baseline on a 2018 benchmark. The durable contribution is that it recognized an underappreciated degree of freedom: one may redesign the representation to expose structural information in a form that a different model family can exploit more efficiently. That intuition aged well even though the original empirical framing did not.

Two aspects of the source paper are especially important in hindsight. First, it emphasized locality preservation. This has remained central because serialization is only useful if nearby tokens continue to provide meaningfully local evidence most of the time. Second, the paper emphasized richer channel capacity than a pure volumetric pipeline could cheaply support. That point is even more relevant in biomolecular AI, where geometry alone is rarely enough. Structural tokens may need atom identity, residue type, electrostatic descriptors, occupancy, or learned latent channels attached to them.

The upgraded perspective thus changes the acceptance story. Instead of asking readers to treat the 2018 experiments as newly competitive, it asks them to see the original method as an early instance of a now-general pattern: ordering is a representational contract between structural data and efficient sequence-native computation. In the remainder of the paper, we make that contract more concrete by pairing the reinterpretation with an executable probe of locality and context-budget exposure.
