\section{Artifact Inventory}

This appendix documents the local files that make the sample reproducible.

\begin{itemize}
\item \texttt{01\_literature\_review.md} through \texttt{04\_research\_review.md}: earlier-stage idea artifacts.
\item \texttt{05\_auto\_review.md}: reviewer-driven refinement notes for the downstream paper story.
\item \texttt{06\_narrative\_report.md}: paper-ready narrative handoff.
\item \texttt{07\_paper\_plan.md}: structure and figure plan.
\item \texttt{results/simulate\_battery\_benchmark.py}: executable stochastic simulator used to generate the benchmark.
\item \texttt{results/demo\_benchmark.json}: computed benchmark data used in the paper.
\item \texttt{results/episode\_metrics.csv}: episode-level metrics for all regimes and policies.
\item \texttt{results/render\_demo\_tables.py}: script that renders the paper tables and figure sources from the JSON artifact.
\item \texttt{code/build\_paper.sh}: end-to-end rebuild wrapper.
\item \texttt{code/generate\_figure\_assets.sh}: figure-export helper.
\item \texttt{code/validate\_artifacts.sh}: PDF and LaTeX validation helper.
\item \texttt{data/dataset\_manifest.json}: simulation-data manifest.
\item \texttt{figure\_assets/}: vector and raster exports for the controller schematic and tradeoff figure.
\item \texttt{review/REVIEW\_OPINION.md}: scored review for the final sample paper.
\end{itemize}

\section{Benchmark Schema}

Each benchmark regime stores the shift family, the battery state, and one row per method. Each method row contains mission return, distance, cost of transport, falls per $100$m, adaptation calls, and adaptation-compute energy. The cost sweep and ablation tables are generated from the same JSON file.

\section{Build Command}

The paper can be rebuilt locally from this folder with:

\begin{verbatim}
cd code && ./build_paper.sh
\end{verbatim}

\section{Package Note}

This example is now packaged to the repo's complete-paper standard: paper source, compiled PDF, code used by the paper, sample data, dedicated figure assets, and a scored review are all included locally. The study remains a lightweight simulator rather than a full robotics benchmark, but the evidence reported in the paper is now generated rather than invented.
