\section{Jefferson and Wilkinson: Access, Usefulness, Protection}

The surviving record suggests that Jefferson and Wilkinson were neither strangers nor intimate friends. They occupied a more typical but harder-to-name early republican relationship: proximity built from usefulness, then tested by crisis.

The relationship is easiest to understand in three stages. The first stage, from 1801 to 1805, is access through usefulness: Wilkinson keeps bringing Jefferson things he can use, whether information, specimens, maps, or western intermediaries. The second stage, centered on 1807, is survival through executive protection: Jefferson's backing matters because Burr makes Wilkinson simultaneously valuable and vulnerable. The third stage, from 1811 to 1818, is memory through selective esteem: Jefferson does not canonize Wilkinson's character, but he continues to credit his services and refuses to join the loudest chorus of suspicion. Read this way, the relationship has a shape. It is political intimacy under pressure, not ordinary friendship.

The early correspondence shows Wilkinson presenting himself as a man worth keeping in Jefferson's circle because he could provide western knowledge, objects, and personnel. In 1802 he sent river-water samples and fossils because they ``may not be unacceptable'' to Jefferson \citep{wilkinson1802}. In 1805 he forwarded a buffalo-hide map and vouched for Captain Amos Stoddard as a man of intelligence, closing with ``perfect respect'' and ``sincere attachment'' \citep{wilkinson1805}. Even the 1804 dinner note is revealing: social access existed, but it was immediately absorbed back into public business through Humboldt, Texas, and Mexico \citep{wilkinson1804jefferson}. The relationship was personal enough for dinner, but the dinner itself seems to have been another venue for utility.

That distinction matters for the paper's use of the phrase \emph{personal relationship}. The letters do show manners, favors, invitations, and affective language. But in early republican elite correspondence those signals do not automatically mean private friendship in the modern sense. They often indicate access, patronage, and mutual placement within a governing circle. Wilkinson's surviving language toward Jefferson is therefore best read as politically personal rather than domestically intimate: close enough for favors, never detached from usefulness.

That pattern became explicit during the Burr crisis. Jefferson's letter of 3 February 1807 is the clearest proof that Wilkinson's survival was tied to presidential need. Jefferson reassured him that ``we ... have not failed to strengthen the public confidence in you'' and promised that he would be ``cordially supported in the line of your duties'' \citep{jefferson1807}. That is not the language of detached neutrality. It is the language of a president who needs his general's credibility to hold while Burr's conspiracy is being narrated to the nation.

Wilkinson understood the asymmetry perfectly. His answer of 20 March 1807 thanked Jefferson for the ``sweet Solace'' of sympathy and described presidential support as ``the warrant of part confidence \& the pledge of future trust'' \citep{wilkinson1807a}. In April he went further and looked ``for Orders \& direction'' concerning ``time \& place'' \citep{wilkinson1807b}. Wilkinson was not writing as a friend seeking consolation. He was writing as a compromised subordinate who needed executive endorsement to remain a usable instrument.

The 1811 exchange sharpens the point. Wilkinson's memoir letter to Jefferson is drenched in persecution and honor. He writes under ``the pressure of my persecutions'' and presents his memoirs as a weapon against public falsehood \citep{wilkinson1811}. Jefferson's answer is warm but carefully framed. He dismisses the ``under my thumb'' slander, says he believes Wilkinson because of his own knowledge of Burr and Wilkinson's conduct, and hopes the general's ``Argosy may ride out'' the storm \citep{jefferson1811}. The letter matters because Jefferson refuses to join the chorus of suspicion. Yet even here the belief is grounded in political history and conduct, not in intimate testimony about character.

Madison's role shows the next step in the same survival chain. Jefferson supplied reassurance, memory, and interpretive cover. Madison later supplied ratification. The distinction matters because Wilkinson did not survive only by remaining believable to Jefferson; he survived when executive sympathy was translated into a restored commission and an approved acquittal. Put differently, Jefferson helped keep Wilkinson politically viable, but Madison helped certify that viability in federal form.

\input{figures/TABLE_themes.tex}
\input{figures/TABLE_phase_modes.tex}

The tables make the stage structure easier to see. Wilkinson's excerpted voice in this corpus carries more defense terms than warmth terms, while Jefferson's language balances warmth with utility and judgment. The phase-by-mode summary then shows why: the Burr crisis collapses almost entirely into protection, while the rehabilitation cluster mixes self-vindication, War Department procedure, testimonial gathering, and official settlement. Wilkinson needed Jefferson more urgently than Jefferson needed Wilkinson personally; Jefferson needed Wilkinson primarily when western policy, Burr, or later retrospective memory made him useful. The 1818 correspondence confirms the shift. Jefferson's estimate of Wilkinson's services remained ``unaltered,'' and Wilkinson treasured the ``approbation \& esteem'' as recognition long denied by the public \citep{jefferson1818,wilkinson1818}. What remained, then, was not intimacy in the ordinary sense. It was a durable politics of remembered service.
